Analyzing Network Transfer Protocols on a Simulated Interplanetary Internet

### 2024-04-09 paper/paper.tex
fix compilation and formatting issues

\usepackage{microtype}
\usepackage{biblatex}

@@

In their 2017 Masters Thesis, Md Monjurul Islam Khan \cite{Khan2017} from University of Manitoba wrote on how 
to create a satellite network
simulation based on ns3. They used a point to point ns3 module to create a network of 2 levels of orbiting 
satellites and fixed points on the earth which acts as uplinks to the network. The paper analyzed TCP 
HighSpeed against TCP Vegas slow start. However this paper stops at expanding the network beyond earth orbit, 
and rather focuses on Earth orbit communication.

SNS3 is a simulation library which uses NS3 as a basis to create a framework which to build simulations for 
geostationary satellites in the european continent. It does not expand to an interplanetary network.
\cite{10.4108/icst.simutools.2014.254631}

For non simulation work done on the interplanetary internet, the leading authority is Consultative Committee for 
Space Data Systems (CCSDS). \cite{CCSDS.org} They maintain and publish the standards for interplanetary internet, including 
the SCPS TP, a reliable transport protocol used for delay tolerant networks when UDP or TCP are not effective. \cite{40565_2004}
Active since 1982 and managed in collaboration with several national and international space organizations. A 
contemporary simulation of a current interplanetary network or earth orbit network would follow the protocols and
standards laid out by CCSDS.

Several public github repos with satellite simulators can be found. 
$\text{https://github.com/szymonwieloch/DTN/tree/master}$
$\text{https://goodingc.github.io/ipn_sim/web-app/}$. Both of these limit themselves to earth orbit simulations.

@@

In a traditional network a majority of the network will happen over a wired connection. This means that the error is relatively 
low. This is no longer the case in the case of satellites sending messages over potentially millions of kilometers. 
Atmosphere in the Earth orbit, solar radiation, and dust cause increased error in sending the message between entities. 
Inverse square law dictates that over these large periods the energy needed to send the message so that it can be 
read at its destination. Even the doppler effect is needed to be accounted for to tune into radio frequencies on the ISS.
High error means that sending a message can no longer be assumed to be sent reliably and has to be compensated for. 
Beyond high error, large distances between entities and the physical limitation that messages cannot travel faster 
than the speed of light. Light, or any other energy takes 8 minutes to get from the Sun to Earth, and thus for a round trip 
will take 16 minutes to complete a simple handshake, given that the messages were sent correctly.

@@

To combat these challenges several protocols have been created. The SCPS TP is a transport protocol which is used on the network 
layer. UDP is used in the lowest reliable connections between nodes, and TCP is used in the most reliable network 
conditions. When a reliable and efficient method of network transportation is needed, then SCPS TP is used. It is a modified 
version of TCP which uses a slow start protocol, reduced handshakes, reliable data transfer with reduced confirmations. Other 
protocols like store and load, and licklider are used to also overcome these problems.

>>>>

\usepackage{microtype}
\usepackage{biblatex}
\usepackage{hyperref}

@@

In their 2017 Masters Thesis, Md Monjurul Islam Khan~\cite{Khan2017} from
University of Manitoba wrote on how to create a satellite network simulation
based on ns3. They used a point to point ns3 module to create a network of 2
levels of orbiting satellites and fixed points on the earth which acts as
uplinks to the network. The paper analyzed TCP HighSpeed against TCP Vegas slow
start. However this paper stops at expanding the network beyond earth orbit, and
rather focuses on Earth orbit communication.

SNS3 is a simulation library which uses NS3 as a basis to create a framework
which to build simulations for geostationary satellites in the European
continent. It does not expand to an interplanetary network~\cite{Puttonen2014}.

For non simulation work done on the interplanetary internet, the leading
authority is Consultative Committee for Space Data Systems
(CCSDS)~\cite{CCSDS.org}. They maintain and publish the standards for
interplanetary internet, including the SCPS TP, a reliable transport protocol
used for delay tolerant networks when UDP or TCP are not
effective.~\cite{Keith2004} Active since 1982 and managed in collaboration with
several national and international space organizations. A contemporary
simulation of a current interplanetary network or earth orbit network would
follow the protocols and standards laid out by CCSDS.

Several public GitHub repositories with satellite simulators can be
found, for instance:
\begin{itemize}
\item\url{https://github.com/szymonwieloch/DTN/tree/master}
\item\url{https://goodingc.github.io/ipn_sim/web-app/}
\end{itemize}
Both of these limit themselves to earth orbit simulations.

@@

In a traditional network a majority of the network will happen over a wired
connection. This means that the error is relatively low. This is no longer the
case in the case of satellites sending messages over potentially millions of
kilometers. Atmosphere in the Earth orbit, solar radiation, and dust cause
increased error in sending the message between entities. Inverse square law
dictates that over these large periods the energy needed to send the message so
that it can be read at its destination. Even the Doppler effect is needed to be
accounted for to tune into radio frequencies on the ISS. High error means that
sending a message can no longer be assumed to be sent reliably and has to be
compensated for. Beyond high error, large distances between entities and the
physical limitation that messages cannot travel faster than the speed of light.
Light, or any other energy takes 8 minutes to get from the Sun to Earth, and
thus for a round trip will take 16 minutes to complete a simple handshake, given
that the messages were sent correctly.

@@

To combat these challenges several protocols have been created. The SCPS TP is a
transport protocol which is used on the network layer. UDP is used in the lowest
reliable connections between nodes, and TCP is used in the most reliable network
conditions. When a reliable and efficient method of network transportation is
needed, then SCPS TP is used. It is a modified version of TCP which uses a slow
start protocol, reduced handshakes, reliable data transfer with reduced
confirmations. Other protocols like store and load, and Licklider are used to
also overcome these problems.
